---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# scrrw

<!-- badges: start -->
<!-- badges: end -->

scrrw fits integrated spatial capture-recapture random walk movement models to estimate wildlife population density and individual movement trajectories

## Installation

You can install scrrw with:

``` r
library(devtools)
install_github("nathan-crum/scrrw")
```

## Simple Random Walk Example

The code below simulates data and fits an integrated spatial capture-recapture simple random walk model

```{r eval = FALSE, echo = TRUE}
library(scrrw)
data = simData_SRW(K.train = 10, K.forecast = 5, det.sd = 1/3, p0 = 0.5, move.sd = 2, A.scale = 2)
mod.fit = scr_lt_Closed_SimpleRW(data = data, density.movement = ~1, g0 = ~1, det.scale = ~1, hSigma = ~-1, hBeta = ~-1)
SCR_lt_Closed_SimpleRW_Abundance(data = mod.fit$data, results = mod.fit$results)
Parameter_Estimates_SimpleRW(data = mod.fit$data, results = mod.fit$results)
FB_SRW = ForwardBackward_Closed_SimpleRW(data = mod.fit$data, par = mod.fit$results$par)
```

```simData_SRW()``` returns a named list, named ```data``` here. ```data``` contains all of the input data needed to fit an integrated spatial capture-recapture simple random walk model using ```scr_lt_Closed_SimpleRW()```.  The objects in ```data``` that are important to pass to ```scr_lt_Closed_SimpleRW()``` for model fitting are:  

* ```y``` is a matrix holding the encounter history
  + The element in row i, column j is 1 if individual i was observed on occasion j; 0 otherwise
* ```y.pix``` is a matrix holding individual's observed locations
  + The element in row i, column j is the grid cell/sub-region number in which individual i was observed on occasion j
  + Grid cell/sub-region numbers correspond to the rows of ```habitat```
* ```detDists``` is a matrix holding individual's distances from survey effort
  + The element in row i, column j is the distance from individual i's observed location on occasion j to the nearest survey trackline or point
* ```y.g0.covar```, ```y.ds.covar```, ```y.hs.covar```, ```y.hb.covar```, and ```y.platform``` are arrays containing any covariate data for the detection model
* ```habitat``` is a matrix with columns named "x", "y", and "Area" describing the grid/sub-regions of the study area  
  + May also contain columns of spatially explicit covariates used to model initial density and movement
* ```distMat``` is a matrix containing the pairwise distances between each pair of grid cells/sub-regions in the study area
* ```detGrid``` is an array describing the detection grid/sub-sub-regions to calculate detection probability
  + Created with ```makeDetGrid()``` 
* ```lines.arr``` is an array holding information about survey effort
  + Contains coordinates for the endpoints and the slope and intercept of each survey trackline
* ```N``` is the number of individuals detected at least once
* ```K``` is the number of sampling occasions
* ```K.prim``` is the number of primary sampling occasions 
  + Must be 1 for the closed population model
* ```SecOcc``` is a vector of the number of secondary sampling occasions in each primary occasion
  + Must sum to ```K```
  + Not used in closed population models
* ```tr_b4_occ``` is a vector of the number of movement transitions before each sampling occasion
  + This can be used to accommodate uneven sampling intervals
* ```nCells``` is the number of grid cells/sub-regions in the study area
  + Must equal the number of rows in ```habitat```
* ```N.states``` is the number of states in the model
  + In the closed population model this must be equivalent to ```nCells```
* ```Area``` is the area of one grid cell/sub-region
* ```nDetCells``` is the number of detection grid cells/sub-sub-regions in the study area
  + Must equal the number of rows in ```detGrid```
* ```detArea``` is the area of one detection grid cell/sub-sub-region
* ```truncDist``` is the truncation distance for the detection model
  + Detection probability is fixed to 0 at detection grid cells that are farther from survey effort than ```truncDist```
* ```mean.ac.move``` is the mean of the maximum pairwise distance between each individual's observed locations
  + Used as an initial value for the movement scale  
* ```log.fact.N``` is a number used in the likelihood calculation
  + Equals ```log(factorial(N))``` or ```sum(log(1:N))```


```scr_lt_Closed_SimpleRW()``` fits a spatial capture-recapture model to the data. The ```density.movement``` argument allows specification of the initial density and movement models using additive functions of any covariates in ```data$habitat```. ```density.movement = ~1``` specifies a model with uniform initial density and no covariates on the movement process. The ```g0```, ```det.scale```, ```hSigma```, and ```hBeta``` arguments allow for specification of the detection models based on covariates in ```detGrid```, ```y.g0.covar```, ```y.ds.covar```, ```y.hs.covar```, and ```y.hb.covar```. ```det.scale``` specifies the model for the scale parameter of the half-normal detection probability, ```hSigma``` and ```hBeta``` specify the models for the parameters in the hazard detection function. If the half-normal model is not used ```det.scale = ~-1``` should be used, and if the hazard model is not used ```hsigma = ~-1, hBeta = ~-1``` should be used.  ```SCR_lt_Closed_SimpleRW_Abundance()``` and ```Parameter_Estimates_SimpleRW()``` return maximum likelihood estimates and 95% confidence intervals for expected and realized density, the movement scale parameter, and detection model parameters.  ```ForwardBackward_Closed_SimpleRW``` returns an array of forward-backward probabilities. ```FB_SRW[[1]][k,j,i]``` is the probability that individual i was in grid cell/sub-region j on occasion k.


## Biased Random Walk Example

The code below simulates data and fits an integrated spatial capture-recapture biased random walk model

```{r eval = FALSE, echo = TRUE}
library(scrrw)
data = simData_BRW(K.train = 10, K.forecast = 5, det.sd = 1/3, p0 = 0.5, move.sd = 2, A.scale = 2)
mod.fit = scr_lt_Closed_BiasedRW(data = data, density.movement = ~1, g0 = ~1, det.scale = ~1, hSigma = ~-1, hBeta = ~-1)
SCR_lt_Closed_BiasedRW_Abundance(data = mod.fit$data, results = mod.fit$results)
Parameter_Estimates_BiasedRW(data = mod.fit$data, results = mod.fit$results)
FB_BRW = ForwardBackward_Closed_BiasedRW(data = mod.fit$data, par = mod.fit$results$par)
```


## Correlated Random Walk Example

The code below simulates data and fits an integrated spatial capture-recapture correlated random walk model

```{r eval = FALSE, echo = TRUE}
library(scrrw)
data = simData_CorrelatedRW(K.train = 10, K.forecast = 5, det.sd = 1/3, p0 = 0.5, move.sd = 2, A.scale = 2)
mod.fit = scr_lt_Closed_CorrelatedRW(data = data, density.movement = ~1, g0 = ~1, det.scale = ~1, hSigma = ~-1, hBeta = ~-1)
SCR_lt_Closed_CorrelatedRW_Abundance(data = mod.fit$data, results = mod.fit$results)
Parameter_Estimates_CorrelatedRW(data = mod.fit$data, results = mod.fit$results)
FB_CRW = ForwardBackward_Closed_CorrelatedRW(data = mod.fit$data, par = mod.fit$results$par)
```
